---
title: "Colleges"
author: "Leo Katz"
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(tidymodels)
library(primer.data)
library(broom)
library(marginaleffects)
```

```{r}
#| cache: true
x <- colleges |>
  select(tuition, grad_rate, selectivity) |>
  filter(tuition > 2) |>
  drop_na() |>
  mutate(selectivity = factor(as.character(selectivity),
                              levels = c("Elite", "Highly Selective",
                                         "Moderately Selective",
                                         "Lowly Selective", "Non-selective")))
```

```{r}
#| cache: true
fit_colleges <- linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + selectivity, data = x)
```

```{r}
#| warning: false
#| message: false
plot_predictions(fit_colleges, condition = c("tuition", "selectivity"), draw = FALSE) |>
  ggplot(aes(x = tuition, y = estimate, color = selectivity)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = selectivity),
              alpha = 0.15, color = NA) +
  geom_line(size = 1) +
  scale_x_continuous(labels = scales::label_dollar(scale = 10000)) +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(
    title = "Predicted Graduation Rate by Tuition and Selectivity",
    subtitle = "Within each selectivity tier, every $10k of tuition adds ~9 percentage points",
    x = "Tuition",
    y = "Predicted Graduation Rate",
    color = "Selectivity",
    fill = "Selectivity",
    caption = "Source: IPEDS 2013 via primer.data"
  ) +
  theme_minimal()
```

There are massive differences in where students choose to enroll, how much they are charged, and their eventual probability of graduating. To assist prospective students in making informed decisions, we analyze the relationship between completion rates, cost of attendance, and school selectivity using 2013 IPEDS records spanning approximately 900 four-year institutions in the United States, which represents our most recent dependable nationwide dataset. However, the 13-year difference between these historical records and our current query presents a notable challenge: because college costs have surged well beyond inflation rates since 2013, the connection between price tag and degree completion may have changed. Our analysis models the graduation rate as a normally distributed outcome that varies linearly with respect to tuition costs and admissions selectivity. According to our estimates, holding selectivity constant, every additional $10,000 in tuition is linked to a roughly 9 percentage point increase in graduation rates (with a 95% confidence interval between 7 and 11 percentage points). Furthermore, Elite universities boast graduation rates that are about 20 percentage points higher than Lowly Selective schools with similar pricing.
